---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

library(dplyr)

## *Load data*
Downloading data from the provided URL
```{r download, echo=TRUE}
if (!file.exists("activity.csv") )
{
  url1 <- 'http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'  
  download.file(url1,destfile='repdata%2Fdata%2Factivity.zip',mode='wb')  
  unzip('repdata%2Fdata%2Factivity.zip')
}
```
## *Read data*
1. Code for reading in the dataset and/or processing the data
```{r read, echo=TRUE}
data <- read.csv("activity.csv")
str(data)
```
## *Steps taken per day*
```{r steps per day, echo=TRUE}
steps_by_day <- aggregate(steps ~ date, data, sum)
summary(steps_by_day)
```

# *Histogram - total steps taken each day*
2.Histogram of the total number of steps taken each day
```{r Histrogram1, echo=TRUE}
hist(steps_by_day$steps, breaks=10, main="Total number of steps per day", xlab="Steps per day")
```

# *Mean & Median - steps taken each day*
3. Mean and median number of steps taken each day
```{r Mean & Meadian, echo=TRUE}
summary(steps_by_day)
```

## *Average daily activity pattern*
```{r Average DAP, echo=TRUE}
steps_by_interval <- aggregate(steps ~ interval, data, mean)
summary(steps_by_interval$steps)
```

# *Plot*
4. Time series plot of the average number of steps taken
```{r Plot-Time Series, echo=TRUE}
plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
```

### *Find row - max of steps*
Row having max steps
```{r Row having max steps, echo=TRUE}
max_steps_row <- which.max(steps_by_interval$steps)
max_steps_row 
```

## *Find interval - max steps*
5. The 5-minute interval that, on average, contains the maximum number of steps
```{r Interval having max steps, echo=TRUE}
steps_by_interval[max_steps_row, ]
```

## *Total no. of missing values in the dataset*
6. Code to describe and show a strategy for imputing missing data
```{r Total missing values, echo=TRUE}
sum(is.na(data$steps))
sum(is.na(data$date))
sum(is.na(data$interval))
```

## *Raplace NAs*
To replace NAs in the dataset
```{r Replace NAs, echo=TRUE}
sum(is.na(steps_by_interval$mean.steps))

data_imputed <- data
for (i in 1:nrow(data_imputed)) {
  if (is.na(data_imputed$steps[i])) {
    interval_value <- data_imputed$interval[i]
    steps_value <- steps_by_interval[
      steps_by_interval$interval == interval_value,]
    data_imputed$steps[i] <- steps_value$steps
  }
}
head(data_imputed)
```
## *Calculate  total number of steps taken each day*
```{r calculate steps per day, echo=TRUE}
df_imputed_steps_by_day <- aggregate(steps ~ date, data_imputed, sum)
head(df_imputed_steps_by_day)
```

## *Histogram*
7. Histogram of the total number of steps taken each day after missing values are imputed
```{r Histogram2, echo=TRUE}
hist(df_imputed_steps_by_day$steps, main="Histogram of total number of steps per day (imputed)", 
     xlab="Total number of steps in a day")
```     
## *To get mean and median of imputed data*
```{r Mean & Median,echo=TRUE}
mean(df_imputed_steps_by_day$steps)

median(df_imputed_steps_by_day$steps)
```

## *get mean and median of data with NA's*
Mean and median with NAs for comparison 
```{r M & M, echo=TRUE}
summary(steps_by_day)
```
## [Mean values stays the same but therer is slight difference in meadian value.]

## *Differences in activity patterns between weekdays and* weekends
```{r DIfference in activity, echo=TRUE}
data_imputed['type_of_day'] <- weekdays(as.Date(data_imputed$date))
data_imputed$type_of_day[data_imputed$type_of_day  %in% c('Saturday','Sunday') ] <- "weekend"
data_imputed$type_of_day[data_imputed$type_of_day != "weekend"] <- "weekday"
```
### **Convert type_of_day from character to factor*
```{r Convert to factor, echo=TRUE}
data_imputed$type_of_day <- as.factor(data_imputed$type_of_day)
```
## *Calculate average steps by interval across all days*
```{r Avg steps, echo=TRUE}
df_imputed_steps_by_interval <- aggregate(steps ~ interval + type_of_day, data_imputed, mean)
```
## Create a plot
8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r Plot2, echo=TRUE}
data_imputed$day <- "weekday"
data_imputed$day[weekdays(as.Date(data_imputed$date), abb=T) %in% c("Sat","Sun")] <- "weekend"

table(data_imputed$day)

##A panel plot containing a time series plot 
library(lattice)
meanStepsPerIntervalNoMissingDay <- aggregate(steps ~ interval + day, data=data_imputed, FUN="mean")
xyplot(steps ~ interval | day, data=meanStepsPerIntervalNoMissingDay, type="l", grid=T, layout=c(1,2), ylab="Number of steps", xlab="5-min. intervals from midnight", main="Average  5-min. activity intervals: Weekdays vs. Weekends")

```
